Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic: adapted to CHOP input format

### 2022-01-26 metanalysis_visualization/visualization.Rmd
add additional outputs from CHOP input data. integrate all the visualizations in one notebook and add new visualizations

Put all the aggregated counts from the different sites together
```{r}
for( i in 1:length( outputFiles )){
  print(i)
  load(paste0( outputDir, outputFiles[i]))
  if( i == 1 ){
    count_icd_all <- count_icd
    patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
    patient_count_psy_period_psy_all <- patient_count_psy_period_psy
    perc_disorder_group_all <- perc_disorder_group
    ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
    bootstrapped_fitted_all <- bootstrapped_fitted_df
    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
    mental_codes_qc_output_all <- mental_codes_qc_output %>%
      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
  } else {
    count_icd_all <- rbind(count_icd_all, count_icd)
    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
    patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
    ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
    bootstrapped_fitted_all <- rbind(bootstrapped_fitted_all, bootstrapped_fitted_df)
    bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
   mental_codes_qc_output <- mental_codes_qc_output %>%
      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all, mental_codes_qc_output)
  }
  rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
     perc_disorder_group,ratios_patients_with_without_psy,
     patient_count_psy_period_psy_clear, table1, bootstrapped_coefficients_df,
     bootstrapped_fitted_df, mental_codes_qc_output)
}
```

@@

```{r}
### add CHOP 
ratios_patients_with_without_psy_chop <- read.delim ( file = "/Users/alba/Desktop/ratios_patients_with_without_psy_chop.txt")

ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy_chop)

ratios_patients_with_without_psy_all %>%
  filter(time_p <= history_end_date &
         time_p >= history_start_date ) %>%
  ggplot( aes(x=time_p, y=percentage, 
              group = siteid, color = siteid))+
  geom_point( aes(shape=siteid, color=siteid) )+
    geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Percentage: patients with psychiatric conditions / total number of patients") +
    xlab("period (months)")
```

@@

### Boxplot showing ratio per month

```{r}
ratios_patients_with_without_psy_all %>%
  ggplot( aes(x=as.factor(time_p), y=ratio)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9,
                aes(shape = siteid)) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Ratio boxplot") +
    xlab("")
```

>>>>

Put all the aggregated counts from the different sites together
```{r}
for( i in 1:length( outputFiles )){
  print(i)
  load(paste0( outputDir, outputFiles[i]))
  if( i == 1 ){
    count_icd_all <- count_icd
    patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
    patient_count_psy_period_psy_all <- patient_count_psy_period_psy
    perc_disorder_group_all <- perc_disorder_group
    ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
    bootstrapped_fitted_all <- bootstrapped_fitted_df
    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
    mental_codes_qc_output_all <- mental_codes_qc_output %>%
      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
  } else {
    count_icd_all <- rbind(count_icd_all, count_icd)
    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
    patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
    ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
    bootstrapped_fitted_all <- rbind(bootstrapped_fitted_all, bootstrapped_fitted_df)
    bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
   mental_codes_qc_output <- mental_codes_qc_output %>%
      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all, mental_codes_qc_output)
  }
  rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
     perc_disorder_group,ratios_patients_with_without_psy,
     patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
     bootstrapped_fitted_df, mental_codes_qc_output)
}
```

@@

```{r}
ratios_patients_with_without_psy_all %>%
  filter(time_p <= history_end_date &
         time_p >= history_start_date ) %>%
  ggplot( aes(x=time_p, y=percentage, 
              group = siteid, color = siteid))+
  geom_point( aes(shape=siteid, color=siteid) )+
    geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Percentage: patients with psychiatric conditions / total number of patients") +
    xlab("period (months)")

ratios_patients_with_without_psy_all %>%
  filter(time_p <= history_end_date &
         time_p >= history_start_date ) %>%
  ggplot( aes(x=time_p, y=ratio, 
              group = siteid, color = siteid))+
    facet_wrap("siteid") +
  geom_point( aes(shape=siteid, color=siteid) ) + 
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Ratio: patients with vs. patients without psychiatric conditions") +
    xlab("period (months)") +
  theme(strip.text = element_text(size = 10),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))
```

@@

### 2022-01-26 psy_peds_3_CHOP.Rmd
add additional outputs from CHOP input data. integrate all the visualizations in one notebook and add new visualizations

@@

@@

# Save the output in a RData file. 
```{r}
site = "CHOP"
patient_count_psy_period_psy$siteid = site 
patient_count_psy_period_psy_clear$siteid = site 
bootstrapped_coefficients_df$siteid = site 
bootstrapped_fitted_df$siteid = site 
ratios_patients_with_without_psy$siteid = site 
quantile_bootstrapped_fitted$siteid = site
perc_disorder_group$siteid = site 
patients_hospitalized_agg_sex_perc$siteid = site

save( ratios_patients_with_without_psy,
      patient_count_psy_period_psy,
      patient_count_psy_period_psy_clear,
      bootstrapped_coefficients_df,
      bootstrapped_fitted_df,
      quantile_bootstrapped_fitted,
      perc_disorder_group,
      patients_hospitalized_agg_sex_perc,
      file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))

>>>>

### Read the ICD10 codes file

The file containing the ICD10 psychiatric related codes is located in
the `public-data` folder of the GitHub repository.

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   dplyr::select( disorder_group, concept_code = ICD10_Code, description )
```

#### QC step: do a summary of the mental health ICD codes present in each site
As output we will generate a table with: the number of observations per code, number of patients diagnosed with each code and the date range in which the code was used.

```{r}
icdCounts_females <- read.csv( "../output/CHOP/Psy-Female-All Ages_ICDlevel.csv") %>%
  pivot_longer(!c(Month, Year), names_to = "concept_code", values_to = "patients") %>%
  mutate(  time_p = ymd(paste( Year, Month, "01", sep="-") ) ) %>%
  select( time_p, concept_code, patients ) %>%
    filter( patients > 0 ) %>%
  group_by( concept_code ) %>%
  summarise( total_females = sum( patients ), 
             min_date = min( as.Date( time_p )) , 
             max_date = max( as.Date( time_p )))

icdCounts_males <- read.csv( "../output/CHOP/Psy-Male-All Ages_ICDlevel.csv") %>%
  pivot_longer(!c(Month, Year), names_to = "concept_code", values_to = "patients") %>%
  mutate(  time_p = ymd(paste( Year, Month, "01", sep="-") ) ) %>%
  select( time_p, concept_code, patients ) %>%
  filter( patients > 0 ) %>%
  group_by( concept_code ) %>%
  summarise( total_males = sum( patients ), 
             min_date = min( as.Date( time_p )) , 
             max_date = max( as.Date( time_p )))

all_codes_ado <- full_join( icdCounts_females, icdCounts_males, by = "concept_code") %>%
  rowwise() %>% 
  mutate( distinct_patients = sum(total_females, total_males, na.rm = TRUE ), 
          distinct_observations = sum(total_females, total_males, na.rm = TRUE ),
          min_date = min( min_date.x, min_date.y, na.rm = TRUE), 
          max_date = max( max_date.x, max_date.y, na.rm = TRUE)) %>%
  select( concept_code, distinct_patients, distinct_observations, min_date, max_date ) %>%
  inner_join( icdCodes )

mental_codes_qc_output  <- all_codes_ado[, c("disorder_group", "description", "concept_code", "distinct_patients", 
                                                    "distinct_observations", "min_date", "max_date")]
mental_codes_qc_output[1:10, ]
```

@@

```{r}
count_icd <- left_join(allCounts, psyCounts) %>%
  mutate(period = ifelse(time_p < as.Date("2020-04-01"),
                         "before_pandemic",
                         "during_pandemic"),
         percentage_psy = count_psy_patients * 100 /count_all,
         siteid = "CHOP") %>%
  select(time_p, count_icd = count_all, period, count_psy = count_psy_patients,
         percentage_psy, siteid)
```

@@

# Save the output in a RData file. 
```{r}
site = "CHOP"
patient_count_psy_period_psy$siteid = site 
patient_count_psy_period_psy_clear$siteid = site 
bootstrapped_coefficients_df$siteid = site 
bootstrapped_fitted_df$siteid = site 
ratios_patients_with_without_psy$siteid = site 
quantile_bootstrapped_fitted$siteid = site
perc_disorder_group$siteid = site 
patients_hospitalized_agg_sex_perc$siteid = site 
mental_codes_qc_output$siteid = site

save( count_icd, 
      ratios_patients_with_without_psy,
      patient_count_psy_period_psy,
      patient_count_psy_period_psy_clear,
      bootstrapped_coefficients_df,
      bootstrapped_fitted_df,
      quantile_bootstrapped_fitted,
      perc_disorder_group,
      patients_hospitalized_agg_sex_perc,
      mental_codes_qc_output,
      file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))
